### 2020-03-05 paper-icns-2020-eda.Rmd
worked on exporting paper to ICNS template

---
title: "Untitled"
author: "RQ"
date: "03/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(gganimate)
```

## Overview

This files puts together the data analytical steps for the pre-processing of the ADSB data for the ICNS paper

One of the ADSB test files was copied from the folder repo to this repo into the data subfolder.
This way the download from OSN could be run in a different sesstion.

## Read In Test Data

```{r}
trjs <- read_csv("./data/adsb_eddf_2019-04-22_0900-1000.csv") %>%
  mutate(time = as_datetime(time))

fact <- read_csv2("../__DATA/EDDF/EDDF_2019_FACT.csv") %>%
  select(FLTID = AP_C_FLTID, REG = AP_C_REG, ADEP = ADEP_ICAO, ADES = ADES_ICAO
         ,MVT_TIME_UTC, BLOCK_TIME_UTC, SCHED_TIME_UTC
         ,ARCTYP
         ,CLASS = AC_CLASS, RWY = AP_C_RWY, STND = AP_C_STND
         ,SRC_PHASE
         ,C40_CROSS_TIME, C40_BEARING,C40_CROSS_LAT,C40_CROSS_LON,C40_CROSS_FL) %>%
  mutate( MVT_TIME_UTC  = lubridate::parse_date_time(MVT_TIME_UTC,   orders="%d.%m.%.Y %T")
         ,BLOCK_TIME_UTC= lubridate::parse_date_time(BLOCK_TIME_UTC, orders="%d.%m.%.Y %T")
         ,SCHED_TIME_UTC= lubridate::parse_date_time(SCHED_TIME_UTC, orders="%d.%m.%.Y %T")
         ,C40_CROSS_TIME= lubridate::parse_date_time(C40_CROSS_TIME, orders="%d.%m.%.Y %T")
         )

acdb <- read_csv("../__DATA/aircraft/aircraft_db.csv") %>%
  mutate( REG = gsub(pattern="-", replacement="", regid)
         ,REG = toupper(REG)
         ,mdl = toupper(mdl))
```
Add icao hex to fact

```{r}
fact2 <- fact %>% left_join(acdb, by = "REG")
dsf  <- fact2 %>% filter(MVT_TIME_UTC >= as.POSIXct("2019-04-22 00:00:00", tz="UTC"),
                        MVT_TIME_UTC <= as.POSIXct("2019-04-23 00:00:00", tz="UTC"))
```

Simple traffic statistics for EDDF

```{r}
traffic_count_per_day <- function(.flts,.start,.end){
  cnt <- .flts %>% 
    mutate(DOF = as.Date(MVT_TIME_UTC)) %>%
    filter(MVT_TIME_UTC >= as.POSIXct(.start, tz="UTC")
          ,MVT_TIME_UTC <= as.POSIXct(.end,   tz="UTC")) %>%
  group_by(DOF, SRC_PHASE) %>%
  summarise(N = n())
return(cnt)
}

start_date <- "2019-04-22 00:00:00"
end_date   <- "2019-05-20 00:00:00"

cnt_fact <- traffic_count_per_day(fact, start_date, end_date)
cnt_fact2<- traffic_count_per_day(fact2, start_date, end_date)
cnt_dsf  <- traffic_count_per_day(dsf  ,"2019-04-22 00:00:00", "2019-04-23 00:00:00")
```

Let's do the traffic count for the study period:

```{r}
cnt_dsf <- cnt_fact2 %>% sample_frac(0.80)    # change this when data crunching ready ########## TOOOOOOOODOOOOOOOOOOOOOOOOO !!!!!!
cnt_dsf  <- cnt_dsf %>% rename(ADSB = N)
cnt_fact <- cnt_fact %>% rename(APDF = N)

cnts <- cnt_fact %>% left_join(cnt_dsf, by=c("DOF","SRC_PHASE"))

ggplot(data = cnts) + 
  geom_col(mapping = aes(x=DOF-0.5, y=APDF, fill=SRC_PHASE), position = "stack", width = 0.3) +
  geom_col(mapping = aes(x=DOF+0.5, y=ADSB, fill=SRC_PHASE), position = "stack", width = 0.3) +
  theme_minimal() +
  theme(legend.position = "none") 
```

Check for what is in the data

```{r}
payload <- trjs %>% group_by(icao24, callsign) %>% 
  summarise(N = n(), GND = sum(onground == TRUE), AIRB = sum(onground == FALSE)) %>% 
  arrange(desc(N)) %>%
  # filter only on gnd
  filter(AIRB >=60) %>%
  # filter overflights
  filter(GND != 0)
```

## check dircetion of flight - arrival or departure

A trjectory is a series of timestamps and lat lon positions, including altitude and other meta data associated with an icao24/callsign tuple (or just one of these).
For each trajectory we can compare the initial and last measurements (i.e. top_n()) and determine whether these are in the air or on the ground.

```{r}
is_arr <- function(.trj, only_guess = TRUE){
  last <- .trj %>% top_n(10, time)
  last = last %>% 
    summarise(ltime = mean(time, na.rm = TRUE), lon_gnd = sum(onground, na.rm=TRUE))
  first = .trj %>% top_n(-10, time)
  first = first %>% 
    summarise(ftime = mean(time, na.rm = TRUE), fon_gnd = sum(onground, na.rm=TRUE))

  out = bind_cols(first, last)
  out <- out %>% 
    mutate(MVT = case_when(
       fon_gnd == 0 & lon_gnd == 0 ~ "OVR"
      ,fon_gnd < lon_gnd ~ "ARR"
      ,lon_gnd < fon_gnd ~ "DEP"
      , TRUE ~ "OTH")
    )

  if(only_guess){out <- out %>% pull(MVT)}  # only provide a flag
  return(out)
}
```

how to handle or extract stationary vehicles or vehicles sitting on ground?
trjs %>% group_by(callsign) %>% summarise(N = n(), ONG = sum(onground), NNA = sum(is.na(onground))) %>% mutate(SITONG = (ONG / N) >=0.95) %>% filter(SITONG == TRUE)

```{r}
# next test sample
ds <- trjs %>% group_by(icao24, callsign) %>% tidyr::nest()
ds <- ds %>% 
  mutate(
    MVT = map(.x=data, .f=~is_arr(.x))
    ,N  = map(.x=data, .f=~nrow(.x))
    ) %>% 
  tidyr::unnest(cols = c(MVT, N))

# write_csv(ds %>% select(icao24, callsign, MVT, N), "./data/EDDF-sample-arr-or-dep.csv")
```

## Load Airport Layout

source osm hack from supporting file. TODO embedd in proper function and add to here.

eddf_ad <- osm_apt(.bb_lonlat = c(8.479385,49.989753,8.616028,50.055596))
eddf_ad

check some of the non movers
trjs %>% group_by(callsign) %>% summarise(N = n(), ONG = sum(onground), NNA = sum(is.na(onground))) %>% mutate(SITONG = (ONG / N) >=0.95) %>% filter(SITONG == TRUE)

stationary <- trjs %>% filter(callsign %in% c("CFG207","CTN413", "FF","FIN7DU","DLH5TP"))
p <- eddf_ad
p <- p + geom_point(data = stationary, mapping=aes(x=lon, y=lat, color=callsign))
p

## Fun Output - Animated Arrival

```{r}
# handpicked flights 
# mt <- trjs %>% filter(callsign %in% c("DLH393","TUI64L","DLH3W","DLH447","LZB437","DLH8HP","DLH465"))

# inspecting ds it seems that N > 500 gives good samples, a few breadcrumbs sit below 100
flts_arr <- ds %>% filter(MVT == "ARR", N > 100)
mt <- trjs %>% filter(callsign %in% (flts_arr %>% pull(callsign)))

viz_stat <- ggplot(data = mt, aes(x=lon, y=lat, color=callsign)) + geom_point() + theme_minimal()
viz_stat

viz_anim <- viz_stat + labs(title = "Time: {frame_time}") + gganimate::transition_time(time)
viz_anim

#save the whole thing
gganimate::anim_save("EDDF_arrs_1h.gif", path = "./data/")
```

## Fun Output - Animated Departures

```{r}
# handpicked flights 
# mt <- trjs %>% filter(callsign %in% c(....))

# inspecting ds it seems that N > 500 gives good samples, a few breadcrumbs sit below 100
flts_dep <- ds %>% filter(MVT == "DEP", N > 100)
mt <- trjs %>% filter(callsign %in% (flts_dep %>% pull(callsign)))

viz_stat <- ggplot(data = mt, aes(x=lon, y=lat, color=callsign)) + geom_point() + theme_minimal()
viz_stat

viz_anim <- viz_stat + labs(title = "Deps - Time: {frame_time}") + gganimate::transition_time(time)
viz_anim

#save the whole thing
gganimate::anim_save("EDDF_deps_1h.gif", path = "./data/")
```

Zoom in on the airport

```{r}
pp <- eddf_ad + geom_point(data = mt, aes(x=lon, y=lat, color=callsign)) + theme_minimal() + xlim(8.475,8.65) + ylim(49.99,50.07) + guides(color=FALSE)
viz_anim2 <- pp + labs(title = "Time: {frame_time}") + gganimate::transition_time(time)
# small size
viz_anim2
# viz_anim3 <- gganimate::animate(viz_anim2, width = 1000, height = 1000)
# gganimate::anim_save("EDDF_deps_1h_zoom.gif", path = "./data/")
```

## Study Departures

```{r}
# first test candidate: 4ca5c8 EIN65W
# 2: 3c65c3 DLH6RL
# 3: 3c4b28 DLH430
# 6: 507c48 UZB232 
md <- flts_dep$data[[2]]

dep_analytic <- function(.trj, .elev_buf = 200){
  md       <- .trj
  elev_buf <- .elev_buf

md2 <- md %>% arrange(time) %>% 
  mutate(lat_prev = lead(lat), lon_prev = lead(lon)
         , lat_d = lat -lat_prev, lon_d = lon - lon_prev)

fst <- md2 %>% filter(time == min(time)) 
fst_pt <-c(fst$lon,fst$lat)

is_next_airb <- function(.airb){ifelse(lag(.airb)=="AIRB", "AIRB", NA)}

md2 <- md2 %>% 
  mutate( id    = row.names(md2)
         ,dist  = geosphere::distHaversine(fst_pt, cbind(lon,lat))
         ,dist2 = dist >= 1
         ,taxi  = ifelse( dist2 == TRUE & onground == TRUE, "TAXI", NA)
         ,lift  = ifelse( is.na(velocity) & is.na(vertrate) & onground == FALSE & baroaltitude <= elev_buf, "LIFT"
                  ,ifelse(is.na(velocity) & is.na(vertrate) & onground == FALSE & is.na(baroaltitude), "LIFT"
                  ,ifelse( velocity > 0   & vertrate > 0    & onground == FALSE & baroaltitude <= elev_buf, "LIFT"
                           , NA) ) )
         ,airb  = ifelse( velocity > 0 & vertrate > 0, "AIRB", NA )
         ,lift2 = is_next_airb(airb)
         )

fst <-  md2 %>% filter(time == min(time)) %>% select(time, lat, lon, id) %>%
  mutate(point = "FIRST", comment = "")

txi <- md2 %>% filter(taxi == "TAXI") %>% filter(time == min(time)) %>%
  select(time, lat, lon, id) %>% 
  mutate(point = "AOBT", comment = "")

tko <- md2 %>% filter(lift == "LIFT" & baroaltitude <= elev_buf) %>%
  select(time, lat, lon, id) %>%
  mutate(point = "ATOT", comment = "") 
if(nrow(tko) > 5){
      tko <- tko[1:5,]
      tko$comment[1]=paste0(" more than 5 LIFTs")}
tko <- tko %>%
  summarise(time = median(time), lat = median(lat), lon = median(lon)
            , id      = paste0(id,      collapse = "+"), point = unique(point)
            , comment = paste0(comment, collapse = " ")
            )

end <- md2 %>% filter(time == max(time)) %>%
  select(time, lat, lon, id) %>%
  mutate(point = "LAST", comment = "")

out <- bind_rows(fst, txi,tko,end)
}

# %>% select(time, lat, lon, lat_prev, lon_prev,lat_d, lon_d, point) 
# md2 %>% select(time, lat, lon, velocity, vertrate, onground)
```

md2[115:125,] %>% select(time, lat, lon, dist, dist2, point)
md2[605:615,] %>% select(time, lat, lon, dist, dist2, velocity, heading, vertrate, onground, point)

```{}
deps <- ds %>% filter(MVT == "DEP") %>%
  mutate(EVENTS = map(.x=data, .f=~dep_analytic(.x)))
```

```{r}
( dep_analytic(flts_dep$data[[9]]) )
```

mydep <- trjs %>% filter(callsign == "EIN65W")

```{r}
dist_pt_to_pt <- function(.trj){
  df <- .trj %>% arrange(time) %>%
    mutate(lat_n = lead(lat), lon_n = lead(lon)
           ,dist = geosphere::distHaversine(cbind(lon, lat), cbind(lon_n,lat_n))
           )
  df$lat_n <- NULL; df$lon_n <- NULL
  return(df)  
}

hold_roll <- function(.trj){
  df <- .trj %>% 
    mutate(
       dist_sum = cumsum(dist)
      ,roll     = ifelse(lead(dist_sum) == dist_sum, "HOLD","ROLL")
      )
  return(df)
}

airborne <- function(.trj){
  df <- .trj %>%
    mutate( airb = ifelse(onground == FALSE, 1, 0)
           ,airb2 =zoo::rollmean(airb, 19, fill=NA)
           ) %>%
    arrange(desc(time)) %>%
    mutate(airb3 = airb2 <1 & !duplicated(airb2 < 1) ) %>%
    arrange(time)
}

prep_trj <- function(.trj){
  trj <- .trj %>% dist_pt_to_pt %>% hold_roll %>% airborne
}

# determine events

first_and_last_hit <- function(.trj){
  first <- .trj %>% filter(time == min(time)) %>% select(time, lat, lon) %>%
    mutate(point = "FIRST")
  last <- .trj %>% filter(time == max(time)) %>% select(time, lat, lon) %>%
    mutate(point = "LAST")
  out <- bind_rows(first, last)
}

block_off <- function(.trj){
  aobt <- .trj %>% filter(dist_sum >= 2)
  aobt <- aobt[1,]  %>% select(time, lat, lon) %>%
    mutate(point = "AOBT")
  return(aobt)
}

take_off_time <- function(.trj){
  atot <- .trj %>% filter(airb3 == TRUE)
  atot <- atot %>% select(time, lat, lon) %>%
    mutate(point = "ATOT")
}

# extract key events

dep_events <- function(.trj){
  fst_lst <- first_and_last_hit(.trj)
  aobt    <- block_off(.trj)
  atot    <- take_off_time(.trj)
  out <- bind_rows(fst_lst, aobt, atot)
}

```

#mydep <- trjs %>% filter(callsign == "EIN65W")
mydep <- trjs %>% filter(callsign == "DLH6RL")
#mydep <- trjs %>% filter(callsign == "UZB232")
#mydep <- trjs %>% filter(callsign == "DLH430")
mydep <- mydep %>% dist_pt_to_pt %>% hold_roll %>% airborne
ggplot(data = mydep, aes(x=time, y=airb2)) + geom_path() + theme_minimal() +
  labs(x="", y="rolling window mean") +
  annotate("text", x = max(mydep$time) - 300, y = 0.5, label=paste0(unique(mydep$callsign)))
# ggsave("./figures/rolling-window-mean.png")

filter(mydep, airb3 == TRUE) %>% select(time, velocity, vertrate, onground, baroaltitude, geoaltitude, dist, dist_sum, roll, airb, airb2,airb3)

mydep[600:615,] %>% select(time, velocity, vertrate, onground, baroaltitude, geoaltitude, dist, dist_sum, roll, airb, airb2,airb3)

mydep <- mydep %>% dist_pt_to_pt %>% hold_roll %>% airborne
ggplot(data = mydep, aes(x=time, y=airb2)) + geom_path()
res   <- dep_events(mydep)
res

ds_dep <- ds %>% filter(MVT == "DEP")
ds_dep <- ds_dep %>% 
  mutate(data2 = map(.x=data, .f=~prep_trj(.x))) %>%
  mutate(out   = map(.x=data2,.f=~dep_events(.x)))

res <- ds_dep %>% select(icao24, callsign, out) %>% 
  tidyr::unnest(cols="out") %>%
  tidyr::pivot_wider(names_from=point, names_prefix="", values_from=time)