### 2020-03-06 x-EDA-01.Rmd
commit after knitr fix

---
title: "Untitled"
author: "Play with OSN and OSM"
date: "20/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggspatial)
library(sf)
library(lubridate)
```

## Overview

This is R Markdown document lists the steps during the exploratory data analysis for the ICNS2020 paper.

Data had been downloaded from OSN (e.g. EDDF).
Chart related data has been extracted from Open Street Map.

## Load data

```{r}
trjs <- read_csv(
  "./data-src/adsb_eddf_2019-04-22_0900-1000.csv"
  ,col_types = cols(
     time   = col_double()
    ,icao24 = col_character()
    ,   lat = col_double()
    ,   lon = col_double()
    ,velocity = col_double()
    , heading = col_double()
    ,vertrate = col_double()
    ,callsign = col_character()
    ,onground = col_logical()
    ,   alert = col_logical()
    ,     spi = col_logical()
    ,  squawk = col_character()
    ,baroaltitude = col_double()
    , geoaltitude = col_double()
    ,lastposupdate = col_double()
    ,  lastcontact = col_double()
    ,  hour = col_double()
  ))
```

TO-DO add osm functions

```{r}
arp_eddf <- geom_point(data = data.frame(lon = 8.56 , lat =50.03 ), aes(x = lon, y = lat), color = "purple", size = 5)
eddf_ad <- osm_apt(.bb_lonlat = c(8.479385,49.989753,8.616028,50.055596))
eddf_ad + arp_eddf
```

arp_eddf <- geom_point(data = data.frame(lon = 8.56 , lat =50.03 ), aes(x = lon, y = lat), color = "purple", size = 5)
+ xlim(8.52, 8.61) + ylim(50, 50.055)

What to do with the trajectories - good candidates for stand identification

lu <- read_csv("../__DATA/EDDF/EDDF_2019_FACT.csv")
lu <- lu %>% filter(SRC_PHASE == "ARR") %>% mutate(TIME = lubridate::dmy_hms(MVT_TIME_UTC, tz="UTC"), DOF = lubridate::date(TIME) ) %>% filter(DOF == as.Date("2019-04-22")) %>% select(FLTID = AP_C_FLTID, REG = AP_C_REG, ADEP_ICAO, ADES_ICAO, MVT_TIME_UTC, BLOCK_TIME_UTC, TIME, DOF, SRC_PHASE)

lu %>% filter(TIME >= lubridate::ymd_hms("2019-04-22 09:00:00", tz="UTC"), TIME <= lubridate::ymd_hms("2019-04-22 10:00:00", tz="UTC"))

```{r}
trjs <- trjs %>%
  mutate(time = as_datetime(time))

payload <- trjs %>% group_by(icao24, callsign) %>% 
  summarise(N = n(), GND = sum(onground == TRUE), AIRB = sum(onground == FALSE)) %>% 
  arrange(desc(N)) %>% ungroup() %>% filter(GND != 0)

# extract some flights with ongrounds
payload %>% filter(GND != 0) %>% sample_n(10)

rq <- trjs %>% right_join(payload, by=c("icao24","callsign"))

ds <- trjs %>% filter(icao24 %in% payload$icao24[10])
```

ggplot(data = ds, mapping = aes(x = lon, y = lat, color = icao24)) + 
  geom_point() +
  xlim(8.52, 8.61) + ylim(50, 50.055)

flts <- c("EIN65W","DLH5RJ","DLH8HP","DLH4YY","DLH695","DLH75U","AAL71","ETD25T","ROT302P")

stat_plot <- ggplot(data = mt, aes(x = lon, y = lat, color=icao24)) + geom_point() + arp_eddf + xlim(8.52, 8.61) + ylim(50, 50.055)
> anim_plot <- stat_plot + gganimate::transition_time(time)
> anim_plot