### 2020-03-10 _paper-2020-ICNS-EDDF.Rmd
clean-up after paper submission

---
title: "paper-2020-ICNS-EDDF"
author: "RQ"
date: "09/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('~/RProjects/paper-2020-ICNS/utility-functions.R')
```

Read in the first day for this study.

```{r}
apt <- "eddf"
dof <- "2019-04-22"

ds  <- load_osn_adsb(apt, dof)

outs<- ds %>% group_by(icao24, callsign) %>%
  summarise(N = n())

# at EDDF there is a good number for ground vehicles transmitting 
# ids FRAxx and LEOSxx --> filter out
outs<- outs %>% filter(N <= 6000)
```

Number of adsb 4D positions per icao24 and callsign

```{r}
g <- ggplot(data = outs, mapping = aes(x=N)) + geom_histogram(bins = 300) +
  theme_minimal() +
  labs(x = "") +
  annotate("text", x = 1200, y = 120, label="count of number of positions \n per icao24/callsign tuple \n at EDDF on 22. April 2019 \n filtered to N <= 6000", hjust = 0)
g
# ggsave("./figures/osn_msg_eddf_20190422.png")
```

Select arrivals and departures from EDDF within study horizon

```{r, message=FALSE}
apt <- "eddf"
dof <- "2019-04-22"

calc_txo_data <- function(.apt, .dof){
  message(paste0("calc txot analytic data for ", .apt, "and day: ",.dof))
apt <- .apt
dof <- .dof
ds  <- load_osn_adsb(apt, dof)

flts_eddf <- read_csv("./data/EDDF-study-flts.csv") 
tdy_deps  <- flts_eddf %>% filter(as.character(DOF) == dof & SRC_PHASE == "DEP")

deps_today<- ds %>% filter(icao24 %in% tdy_deps$icao24) %>%
  mutate(DIST_ARP = geosphere::distHaversine(c(8.5705,50.04), cbind(lon, lat) ) ) %>%
  group_by(icao24, callsign) %>%
  filter(time == min(time)) %>%
  filter(DIST_ARP <= 4000) %>%
  select(icao24, callsign) %>% ungroup()

deps <- ds %>% inner_join(deps_today) %>% group_by(icao24, callsign) %>%
  mutate(nx_lon = lag(lon), nx_lat = lag(lat)
         ,DIST = geosphere::distHaversine(cbind(lon, lat), cbind(nx_lon, nx_lat))
         ,DIST = ifelse(is.na(DIST), 0, DIST)
         ,DIST2= cumsum(DIST)
         ,DIST3= zoo::rollmean(DIST2, 13, fill=0 )
         )

hits <- deps %>% summarise(N = n())
  # remove many hit - think about this

first <- deps %>%
  filter(time == min(time)) %>%
  mutate(EVENT = "FIRST")

taxi <- deps %>%
  filter(DIST2 > 2 & DIST3 > 3) %>%
  filter(time == min(time)) %>%
  mutate(EVENT = "AOBT")

airb <- deps %>% 
  mutate( airb  = ifelse(onground == FALSE, 1, 0)
        ,airb2 = zoo::rollmean(airb, 19, fill=NA)
  ) %>%
    arrange(desc(time)) %>%
    mutate(airb3 = airb2 <1 & !duplicated(airb2 < 1) ) %>%
    arrange(time) %>%
  filter(airb3 == TRUE) %>% mutate(EVENT = "ATOT")

comb <- bind_rows(first, taxi, airb) %>%
  select(time, icao24, callsign, EVENT) %>%
  inner_join(hits, by = c("icao24","callsign")) %>%
  mutate(DOF = dof)
}
```

```{r}
# rq <- calc_txo_data(apt, dof)
apt <- "eddf"
dofs <- c( "2019-04-22","2019-04-23","2019-04-24","2019-04-25","2019-04-26","2019-04-27","2019-04-28"
          ,"2019-04-29","2019-04-30","2019-05-01","2019-05-02","2019-05-03","2019-05-04","2019-05-05"
          ,"2019-05-06","2019-05-07","2019-05-08","2019-05-09","2019-05-10","2019-05-11","2019-05-12"
          ,"2019-05-13","2019-05-14","2019-05-15","2019-05-16","2019-05-17","2019-05-18","2019-05-19")
# iterator over apt and dofs
it  <- expand.grid(APT = as.character(apt), DOF = dofs)

tx_eddf <- it %>% 
  purrr::pmap_dfr(.l=., .f=~calc_txo_data(.apt=..1, .dof=..2))

write_csv(rq, "./data/EDDF_analytic_data.csv")
```

couple results and apdf

```{r}
cols_flts <- cols(
  DOF = col_date(format = ""),
  SRC_PHASE = col_character(),
  AP_C_FLTID = col_character(),
  REG = col_character(),
  ARCTYP = col_character(),
  AC_CLASS = col_character(),
  AP_C_RWY = col_character(),
  AP_C_STND = col_character(),
  icao24 = col_character(),
  model = col_character(),
  typecode = col_character()
)
tx_eddf   <- read_csv("./data/EDDF_analytic_data.csv")
flts_eddf <- read_csv("./data/EDDF-study-flts.csv", col_types = cols_flts) 
icao_and_reg <- flts_eddf %>% select(icao24, REG) %>% unique() %>% na.omit()
gate_and_rwy <- flts_eddf %>% select(REG, AP_C_RWY, AP_C_STND, DOF) %>% na.omit()

tx_eddf <- tx_eddf %>% left_join(icao_and_reg, by="icao24") %>%
  filter(N >= 100)
tx_eddf <- tx_eddf %>% group_by(icao24, REG, DOF) %>% 
  arrange(icao24) %>% 
  pivot_wider(names_from = EVENT, values_from = time)

#tx_eddf <- tx_eddf %>% left_join(gate_and_rwy, by=c("DOF","REG"))
# shortcut here for paper

out <- tx_eddf %>% ungroup() %>%
  mutate(TXOT = difftime(ATOT,AOBT,units = "mins")) %>%
  filter(TXOT > 0, TXOT <=180)

ref <- out %>% summarise(REF = quantile(TXOT, probs = 0.2))

out <- out %>% mutate(REF = ref$REF
                      ,ADD_TXOT = TXOT - REF)
add_txot <- out %>% summarise(TXOT = sum(TXOT, na.rm = TRUE), REF = sum(REF, na.rm=TRUE)
                         , ADD_TXOT = sum(ADD_TXOT, na.rm = TRUE)
                         , N = nrow(out)
                         , N_NA = sum(is.na(ADD_TXOT))
                         ) %>%
  mutate(AVG_ADD_TXOT = ADD_TXOT / (2* N))

write_csv(add_txot, "./data/EDDF_analytic_txot.csv")
```

rq <- tx_eddf %>% group_by(icao24, REG, DOF)

df <- .trj %>%
    mutate( airb = ifelse(onground == FALSE, 1, 0)
           ,airb2 =zoo::rollmean(airb, 19, fill=NA)
           ) %>%
    arrange(desc(time)) %>%
    mutate(airb3 = airb2 <1 & !duplicated(airb2 < 1) ) %>%
    arrange(time)

```{r}
is_close_and_low <- function(.ds, .arp=c(8.5705,50.04)){
# arrival: later timestamps near airport and low altitude 
# departure: early timestamps near airport and low altitude
  ds <- .ds %>% 
    arrange(time)%>%
    mutate(DIST_ARP = geosphere::distHaversine(.arp, cbind(lon, lat) ) ) %>% 
    group_by(icao24, callsign)
# check distance and alt
  first <- ds %>% top_n(15, time) %>%
    summarise( time     = mean(time    , na.rm = TRUE)
              ,DIST_ARP = mean(DIST_ARP, na.rm = TRUE)
              ,ALT      = mean(geoaltitude, na.rm = TRUE)
              ,ONG      = mean(sum(onground, na.rm = TRUE))
              ) %>% 
    mutate(EVENT = "FIRST")
  last  <- ds %>% arrange(time) %>% top_n(-15, time) %>%
    summarise( time     = mean(time    , na.rm = TRUE)
              ,DIST_ARP = mean(DIST_ARP, na.rm = TRUE)
              ,ALT      = mean(geoaltitude, na.rm = TRUE)
              ,ONG      = mean(sum(onground, na.rm = TRUE))
              ) %>% 
    mutate(EVENT = "LAST")
  return(bind_rows(first, last))
}
```

# remove overflights -------------------------
  ovr <- ds %>% group_by(icao24, callsign) %>% 
    summarise( GEO_ALT_MIN = min(geoaltitude, na.rm = TRUE)
              ,GEO_ALT_NA  = sum(is.na(geoaltitude))
              , N          = n()) %>%
    filter(GEO_ALT_MIN >=5000, GEO_ALT_NA / N < 0.4) %>%
    ungroup %>% select(icao24, callsign)
  ds <- ds %>% anti_join(ovr, by=c("icao24","callsign"))

pick_daily_deps <- function(.ds, .dof, .flts){
  flts <- .flts %>% filter(DOF == .dof)
  flts_deps <- flts %>% filter(SRC_PHASE == "DEP")

  ds <- .ds %>% filter(icao24 %in% flts_deps$icao24)
}