### 2020-03-10 index.Rmd

---
title: ICNS2020 - paper - Operational Performance Analyses with Community Collected Open Data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

some pretty cool stuff I am going to write about the paper.

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.